Manuscript draft

### 2019-08-28 manuscript/manuscript draft.Rmd
updated manuscript draft

@@

Related to this, we find that climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes. Although this could in part be related to a lack of standardisation in methodology across studies, as discussed above, it could also indicate that factors other than climate have significant influence over subsidiary fluxes. This is consisent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes. We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.

Past studies have differed in their conclusions regarding the relationship between productivity and latitude. Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions.

There is strong evidence from our analyses that this increase in productivity towards the tropical regions is driven primarily by temperature-related effects. We find that temperature-related climate variables explain the highest proportion of variability in productivity.

These conclusions are supported by our analyses of the influence of climate within the growing season. We find much weaker correlations between climate variables and productivity when only growing season months are considered. There is evidence for a positive correlation between productivity and temperature for ANPP and ANPP~foliage~, suggesting that there is still a small positive effect of air temperature on productivity independent of growing season, however this effect is not significant for the remaining fluxes. This suggests that the primary positive effect of temperature on productivity is through the lengthening of the growing season in regions with higher MAT. However, we note that length of growing season is a less good predictor of productivity than MAT. This may be an artefact of imprecise estimation of length of growing season, and with more precise climate data on finer time-scales it is likely that the predictive ability of length of growing season would increase, however it does indicate that there are aspects of productivity explained by MAT which length of growing season alone cannot explain. It is hard to attribute mechanisms through such coarse-scale data as this. Future studies.......

The majority of studies to date have focused on MAT and MAP as two key climate variables for which data is easily obtained. We find that MAP is not a significant predictor of productivity, but does show a strong interaction with temperature. This indicates that the main mechanism through with precipitation influences productivity is in a secondary capacity, with its overall influence being mediated by the effects of temperature. This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to understand patterns of productivity. This is likely to be significant in the future, as the effects of a warming climate will be mediated through changes in precipitation regime.

This studies considers the relationship between productivity and climate on a coarse scale. Because of the nature of the ForC database, we were unable to access site-level MAT and MAP data for a number of sites, and lacked site-level data for other climate variables for all sites. As a result the climate data we used was primarily multi-year averages of annual climate measures from global databases. While the quality of these data is high, we cannot expect that it captures detailed local variation in climate. Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT. With high-quality stand-level data for other climate variables we might find stronger predictors. Similarly, with accurate stand-level measurements of growing season length, we would expect to record a stronger correlation between growing season length and productivity.

*The relationship between productivity and latitude*

We are no longer explicity testing whether climate explains a higher proportion of variability in major fluxes, so mention this, and the implications of it in the discussion.

Past studies have differed in their conclusions regarding the relationship between productivity and latitude. Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions. While Huston and Wolverton (2009) found no significant relationship between ANPP and latitude, we find a significant negative relationship between ANPP and latitude, with the highest levels of ANPP in the tropical regions.

*Challenges in estimating NPP*  
The impact of excluding certain components of NPP on this analysis is unclear, primarily because these components are poorly understood and rarely measured [@clark_net_2001]. Do we want to include a section assessing how much this may or may not impact results?  
1. Hard to measure BNPP because it often isn't sampled deep enough; impact on ratio tests?  
2. Unclear if there are latitudinal patterns e.g. of herbivory/VOCs  
  + Herbivory can be an important component of total NPP so this may be important to account for; likely to be higher in the tropics?

>>>>

Past studies have differed in their conclusions regarding the relationship between productivity and latitude. Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions. There is strong evidence from our analyses that this increase in productivity towards the tropical regions is driven primarily by temperature-related effects. We find that temperature-related climate variables explain the highest proportion of variability in productivity. In contrast, we find that MAP is not a significant predictor of productivity, but does show a strong interaction with temperature. This indicates that the main mechanism through with precipitation influences productivity is in a secondary capacity, with its overall influence being mediated by the effects of temperature. This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to understand patterns of productivity. This is likely to be significant in the future, as the effects of a warming climate will be mediated through changes in precipitation regime.

These conclusions are supported by our analyses of the influence of climate within the growing season. We find much weaker correlations between climate variables and productivity when only growing season months are considered. There is evidence for a positive correlation between productivity and temperature for ANPP and ANPP~foliage~, suggesting that there is still a small positive effect of air temperature on productivity independent of growing season, however this effect is not significant for the remaining fluxes. This suggests that the primary positive effect of temperature on productivity is through the lengthening of the growing season in regions with higher MAT. However, we note that length of growing season is a less good predictor of productivity than MAT. As discussed above, this may be an artefact of imprecise estimation of length of growing season, however it does indicate that there are aspects of productivity explained by MAT which length of growing season alone cannot explain.

@@

Related to this, we find that climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes. Although this could in part be related to a lack of standardisation in methodology across studies, as discussed above, it could also indicate that factors other than climate have significant influence over subsidiary fluxes. This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes. We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.
